% !TeX spellcheck = en_UK
% \let\accentvec\vec              
\documentclass[runningheads,11pt]{llncs}
%\let\spvec\vec \let\vec\accentvec
\newcommand\hmmax{0}
\newcommand\bmmax{0}

\usepackage{amssymb,amsmath}
%\let\vec\spvec
%\usepackage{lmodern}
\usepackage{newtxmath,newtxtext}
\usepackage[T1]{fontenc}
% \def\vec#1{\mathchoice{\mbox{\mathbf$\displaystyle#1$}}
% {\mbox{\mathbf$\textstyle#1$}} {\mbox{\mathbf$\scriptstyle#1$}}
% {\mbox{\mathbf$\scriptscriptstyle#1$}}}

\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
\DeclareFontShape{U}{mathx}{m}{n}{<-> mathx10}{}
\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
\DeclareMathAccent{\widebar}{0}{mathx}{"73}

\usepackage{soulutf8} \soulregister\cite7 \soulregister\ref7
\soulregister\pageref7 \usepackage{hyperref}
\usepackage[color=yellow]{todonotes} \hypersetup{final} \usepackage{mathrsfs}
\usepackage[advantage,asymptotics,adversary,sets,keys,ff,lambda,primitives,events,operators,probability,logic,mm,complexity]{cryptocode}

\usepackage{cite} 
\usepackage{booktabs}
\usepackage{paralist}
\usepackage[innerleftmargin=5pt,innerrightmargin=5pt]{mdframed}
\usepackage{caption}
\captionsetup{belowskip=0pt}
\usepackage{bm}
\usepackage{url}
%\usepackage{dirtytalk}
\newcommand{\say}[1]{\emph{``#1''}}
\usepackage[margin=0.7in,a4paper]{geometry}
\usepackage[normalem]{ulem}
\usepackage{dashbox}
\newcommand{\dboxed[1]}{\dbox{\ensuremath{#1}}}
\usepackage{hyperref}
\usepackage[capitalise]{cleveref}

%\usepackage{mathptmx}

\include{macros}

\title{On PIOP-based zkSNARKs}

\author{} 
%\iflncs{
\institute{} 

\allowdisplaybreaks

\begin{document} \sloppy \maketitle

\begin{abstract}
  In this paper we investigate properties of zkSNARKs obtained by
  compiling a PIOP proof using a polynomial commitment scheme. The question we
  try to answer is \say{What polynomial commitment's properties propagate to
    the resulting zkSNARK?}. The properties we focus on are:
  \begin{compactenum}
  \item simulation extractability,
  \item SRS updatability,
  \item SRS-updatable simulation extractability,
  \item subversion zero knowledge.
  \end{compactenum}
  The research hypothesis is \say{A NIZK obtained from a simulation extractable /
    SRS-updatable / SRS updatable SE / subversion zero knowledge polynomial
    commitment is simulation extractable /
    SRS-updatable / SRS updatable SE / subversion zero knowledge.} To be able to
  show the hypothesis we need to solve a number of problems
  \begin{compactenum}
  \item Neither simulation extractability, SRS-updatability, SRS-updatable
    simulation extractability, nor subversion zero knowledge have been defined
    for a polynomial commitment scheme. Another contribution of the paper is
    defining these properties. 
  \item Similarly, there is no definition for SRS-updatable simulation
    extractable NIZKs.
  \item The polynomial IOP is defined very generally, cf.~\cref{def:piop}, what
    makes showing generic properties difficult. The paper would propose tighter
    definitions which would emphasize more the structure of PIOP,
    but would not narrow a class of possible (from the practical point of view)
    PIOPs too much, cf.~\cref{def:wepiop,def:sdwepiop}.
  \end{compactenum}
  
\end{abstract}

\paragraph{Polynomial commitment scheme.}
\label{sec:poly_com}
In the polynomial commitment scheme $\PCOM = (\kgen, \com, \open, \verify)$ the
prover $\prover$ convinces the verifier $\verifier$ that some polynomial $\p{f}$
which $\prover$ committed to evaluates to $s$ at some point $z$ chosen by
$\verifier$.  The key generation algorithm $\kgen$ takes as input a security
parameter $\secparam$ and a parameter $\maxdeg$ which determines the maximal
degree of the committed polynomial. We assume that $\maxdeg$ can be read from
the output SRS.
  
We emphasize the following properties of a secure polynomial commitment
$\PCOM$:
\begin{description}
\item[Evaluation binding:] A $\ppt$ adversary $\adv$ which outputs a commitment
  $\vec{c}$ and evaluation points $\vec{z}$ has at most negligible chances to
  open the commitment to two different evaluations $\vec{s}, \vec{s'}$. That is,
  let $k \in \NN$ be the number of committed polynomials, $l \in \NN$ number of
  evaluation points, $\vec{c} \in \GRP^k$ be the commitments,
  $\vec{z} \in \FF_p^l$ be the arguments the polynomials are evaluated at,
  $\vec{s},\vec{s}' \in \FF_p^k$ the evaluations, and
  $\vec{o},\vec{o}' \in \FF_p^l$ be the commitment openings. Then for every
  $\ppt$ adversary $\adv$
	\[
		\Pr
			\left[
			\begin{aligned}
				& \verify(\srs, \vec{c}, \vec{z}, \vec{s}, \vec{o}) = 1,  \\ 
				& \verify(\srs, \vec{c}, \vec{z}, \vec{s}', \vec{o}') = 1, \\
				& \vec{s} \neq \vec{s}'
			\end{aligned}
			\,\left|\,\vphantom{\begin{aligned}
                  & \\
                  & \\
                  &
                \end{aligned}}
			\begin{aligned}
				& \srs \gets \kgen(\secparam, \maxdeg),\\
				& (\vec{c}, \vec{z}, \vec{s}, \vec{s}', \vec{o}, \vec{o}') \gets \adv(\srs)
			\end{aligned}
			\right.\right] \leq \negl\,.
	\]

\end{description}
	

\begin{description}
\item[Commitment of knowledge]  For every $\ppt$ adversary $\adv$ who produces
  commitment $c$, evaluation point $z$, evaluation $s$ and opening $o$ there
  exists a $\ppt$ extractor $\ext$ such that
\[
  \Pr \left[
    \begin{aligned}
      & \p{f} = \ext_\adv(\srs, c; r),\\
      & c = \com(\srs, \p{f}),\\
      & \verify(\srs, c, z, s, o) = 1
    \end{aligned}
    \,\left|\,
      \vphantom{
        \begin{aligned}
          & \\
          & \\
          &
        \end{aligned}
        }
    \begin{aligned}
      & \srs \gets \kgen(\secparam, \maxdeg),\\
      & (c, z, s, o) \gets \adv(\srs; r)
    \end{aligned}
  \right.\right]
  \geq 1 - \epsk(\secpar).
\]
In that case we say that $\PCOM$ is $\epsk$-knowledge.
\end{description}
Intuitively when a commitment scheme is a commitment of knowledge then if an
adversary produces a (valid) commitment $c$, which it can open, then it also
knows the underlying polynomial $\p{f}$ which commits to that value.

\begin{definition}[Simulation extractable polynomial commitment]
  \label{def:sepcom}
  Let $\PCOM$ be a polynomial commitment scheme, let $\oracles$ be a simulator
  oracle which on input
  \begin{description}
  \item [$(\msg{commit}, f)$:] returns commitment $c = \com(f)$ and adds
    $(f, c)$ to list $Q$; Optionally the oracle may be queried on a pair
    $(\msg{rand}, d)$ instead of $f$. In that case the oracle picks a random
    polynomial of degree $d$ and outputs its commitment.
  \item[$(\msg{evaluate}, c, x)$:] returns $f(x)$ where $f$ is a polynomial
    which commitment is $c$ and $(f, c) \in Q$
  \item[$(\msg{open}, c, x, y)$:] returns an opening assuring that for a
    polynomial $f$, such that $c = \com(f)$, $f(x) = y$. If $c$ is not a
    commitment output by the oracle or $f(x) \neq y$, then ignore.
  \end{description}
  We say that $\PCOM$ is \emph{simulation extractable} if for any $\ppt$
  adversary $\adv$ with oracle access to $\oracles$ there exists extractor
  $\ext$ such that
  \[
    \Pr\left[
      \begin{aligned}
        & (\deg{f} > \maxdeg) \lor\\
        & (c \not\in \image(\com(f)))
        \end{aligned}
        \,\left|\,
          \begin{aligned}
          & \pp \sample \sgen(\secparam),
           \srs \sample \kgen(\pp, \maxdeg),\\
           & c \gets \adv^{\oracles}(\pp, \srs; r),
           (\cdot, c) \not\in Q, 
           f \gets \ext_{\adv}(\secparam, \pp, c, r)
        \end{aligned}
      \right.  \right] \leq \negl.
  \]
  \michals{23.04}{Can $\ext$ ask $\adv$ to give evaluations of the committed
    polynomial? That is how $\ext$ in a proof system works -- it evaluates
    polynomials submitted by the adversary on multiple challenges.}
  
\end{definition}


\begin{definition}[Polynomial IOP,~\cite{EPRINT:Szepieniec20}]
  \label{def:piop}
  Let $\REL$ be an indexed relation with a corresponding language $\LANG$, $\FF$
  some finite field, and $\maxdeg$ a degree bound and $\noofp$ a parameter. A
  \emph{polynomial IOP for $\REL$ with degree bound $\maxdeg$} is a pair of
  interactive machines $\prover, \verifier$ such that
\begin{itemize}
\item $(\prover, \verifier)$ is an interactive proof for $\LANG$ with $r$ rounds
  and soundness error $\epss$;
\item $\prover$ sends to $\verifier$ polynomials $f_i \in \FF[X]$,
  $i \in \range{1}{\noofp}$, of degree at most $\maxdeg$;
\item $\verifier$ is an oracle machine with access to a list of oracles, which
  contains one oracle for each polynomial it has received from the prover;
\item When an oracle associated with a polynomial $f_i(X)$ is queried on a point
  $z_j \in \FF$, the oracle responds with the value $f_i(z_j)$; 
\item $\verifier$ sends challenges $\alpha_k \in \FF$ to $\prover$;
\item $\verifier$ is public coin.
\end{itemize}
\end{definition}

\michals{28.04}{Add preprocessing and zero knowledge}

\begin{definition}[Witness encoding PIOP (WEPIOP)]
  \label{def:wepiop}
  Let $\PS$ be a PIOP.  We say that $\PS$ is \emph{witness encoding} if there is
  a function $\decode$ and set $\encset \in [\noofp]$ such that for any
  $(\inp, \wit) \in \REL$ and polynomials $\smallset{f_i}_{i \in [\noofp]}$ sent by an
  honest prover, $\decode(\smallset{f_i}_{i \in \encset}) = \wit$. We call $\encset$ the
  \emph{encoding set}.
\end{definition}
In other words, PIOP is witness encoding if for any valid proof for a statement
$\inp$ in the language, the corresponding witness can be read from the
polynomial coefficients. We note that this is the case for virtually all
PIOPs. \michals{28.04}{Check!}]

\begin{definition}[Somehow deterministic WEPIOP]
  \label{def:sdwepiop}
  Let $\PS$ be a WEPIOP for $\REL$. For each polynomial $f_i$ sent by the prover
  denote by $A_i$ the set of challenges sent by the verifier and by $F_i$ the
  set of polynomials sent by the prover \emph{before} the prover sends
  $f_i$. Let $\encset$ be an encoding set. We say that $\PS$ is \emph{somehow
    deterministic} (SD) if for any $(\inp, \wit) \in \REL$, polynomials
  $\smallset{f_i}_{i \in [\noofp]}$ send by the prover, and encoding set
  $\encset \neq [\noofp]$ each polynomial
  $f_j \in \smallset{f_i}_{i \in [\noofp] \setminus \encset}$ is determined by
  \begin{itemize}
    \item polynomials $F_j$, and
    \item the verifier's challenges $A_i$, and
    \item the witness $\wit$.
  \end{itemize}
\end{definition}
Intuitively, we say that WEPIOP is somehow deterministic if the only
non-deterministic messages send by the prover are polynomials encoding the
witness, and all other messages are determined by the previous one, witness, and
verifier's challenges.

\begin{lemma}
  Let $\PS = (\PS.\prover, \PS.\verifier)$ be a PIOP with knowledge soundness error
  $\epsks$ where the prover sends up to $\noofp$ polynomials. Let $\PCOM$ be a
  knowledge polynomial commitment scheme with extraction error $\epsext$. Let
  $\PSc = (\PSc.\prover, \PSc.\verifier)$ be a proof system such that
  \begin{description}
  \item[$\PSc.\prover$] acts as $\PS.\prover$, except
    \begin{itemize}
    \item when $\PS.\prover$ sets up a polynomial oracle $\oracleo_f$,
      $\PSc.\prover$ sends commitment $c = \com(f)$;
    \item when $\PSc.\verifier$ asks $\PSc.\prover$ to open a commitment
      $c = \com(f)$ at $z$ it returns $f(z)$.
  \end{itemize}
  \item[$\PSc.\verifier$] acts as $\PS.\verifier$, except when $\PS.\verifier$
    asks oracle $\oracleo_f$ for an evaluation of $f$ at $z$, $\PSc.\verifier$
    sends $z$ to $\PSc.\prover$ and gets $f(z)$ in return.
  \end{description}
  Then $\PSc$ is sound with soundness error at most $\epsks + \noofp \cdot \epsext$.
\end{lemma}
\begin{proof}
  % Let $\adv$ be an adversary that breaks knowledge soundness of $\PSc$ with
  % probability $\eps$ greater than $\epsks + \noofp \cdot \epsext$. We build
  % $\bdv$ that breaks knowledge soundness of $\PS$ with probability greater than
  % $\epsks$.

  % $\bdv$ starts by picking an SRS $\srs$ for the polynomial commitment scheme
  % $\PCOM$ and provides $\srs$ to $\adv$. Let $\inp$ be the instance $\adv$
  % creates a fraudulent proof for. Each time $\adv$ sends a commitment $c$, then
  % $\bdv$ uses the extractor $\ext$ to extract polynomial $f$ from $c$ and sends
  % $f$ to the newly set oracle $\oracleo_f$.  \michals{5.5}{how much we can
  %   generalize here?  I.e.~could we have a single proof that would cover
  %   algebraic adversaries that returns with $c$ also a vector of group elements'
  %   logarithms; or commitment schemes where there is an extraction key $k$
  %   allowing to extract $f$ from $c$?}  Each time $\PS.\verifier$ sends a
  % challenge $\alpha$, $\bdv$ passes $\alpha$ to $\adv$ as $\PSc.\verifier$'s
  % challenge and sends $\adv$'s response back to $\PS.\verifier$. Eventually,
  % $\adv$ finalizes its proof for $\inp$. Note that if $\PSc.\verifier$ accepts
  % the proof provided by $\adv$, then $\PS.\verifier$ accepts the proof provided
  % by $\bdv$. Probability that $\bdv$ \emph{fails} is upper-bounded by two
  % events. One is that $\adv$ fails to provide a convincing proof such that no
  % extractor can output a corresponding witness. Probability of this event is
  % upper-bounded by $(1 - \eps)$. Furthermore, $\bdv$ may fail to extract some
  % of the polynomials which $\adv$ made commitments for. Since $\adv$ commits to
  % $\noofp$ polynomials and $\bdv$ fails to extract each of them with probability
  % at most $\epsext$, by the union bound $\bdv$ fails to extract \emph{some} of
  % the polynomials with probability upper-bounded by $\noofp \cdot
  % \epsext$. Hence the probability of $\bdv$'s failure is upper-bounded by
  % \[
  %   (1 - \eps) + \noofp \cdot \epsext.
  % \]
  % Hence, for $\eps > \epsks + \noofp \cdot \epsext$ adversary $\bdv$ wins with probability at least
  % \[
  %   \epsks + \noofp \cdot \epsext - \noofp \cdot \epsext  = \epsks.
  % \]
  % \qed
\end{proof}


\bibliographystyle{alpha}
\bibliography{cryptobib/abbrev1,cryptobib/crypto}

\end{document}